Arrows for Parallel Computations

### 2017-05-18 paper/content/main.tex
it compiles

%% add a tweak to polycode, similar to literate in lstlistings, but uglier
%% ^^^^ you want this. and similar for ***, +++, etc.
%% a hack for list comprehensions' <-, typeset it as <--
%%%%%format <--        = "\in "
%fornat ` = "\`"

@@

@@

\subfloat[Arrow class definition][]{%

@@

@@

\>[3]{}(\mathbin{>\!>\!>})\mathbin{::}\Varid{arr}\;\Varid{a}\;\Varid{b}\to \Varid{arr}\;\Varid{b}\;\Varid{c}\to \Varid{arr}\;\Varid{a}\;\Varid{c}{}\<[E]%

@@

% caption{Arrow class definition}
}%\hfill
\subfloat[The arrow type class definition on the left with schematic depiction of its combinators on the right][]{%
	\begin{center}

@@

	\end{center}
}%\hfill
%\caption{The arrow type class definition on the left with schematic depiction of its combinators on the right}

@@

\lstinline{arr} is used to lift an ordinary function to an arrow type, similarly to the monadic \lstinline{return}. The \lstinline{>>>} operator is analogous to the monadic composition  \lstinline{>>=} and combines two arrows \inlinecode{arr a b} and \inlinecode{arr b c} by "wiring" the outputs of the first to the inputs to the second to get a new arrow \inlinecode{arr a c}. Lastly, the \lstinline{first} operator  takes the input arrow from \lstinline{b} to \lstinline{c} and converts it into an arrow on pairs with the second argument untouched. It allows us to to save input across arrows.

@@

The most prominent instances of this interface are regular functions \lstinline{(->)} (Fig.~\ref{fig:arrowfn}),

@@

\>[9]{}\Varid{f}\mathbin{>\!>\!>}\Varid{g}\mathrel{=}\Varid{g}\mathbin{\circ}\Varid{f}{}\<[E]%

@@

\>[9]{}\Varid{f}\mathbin{>\!>\!>}\Varid{g}\mathrel{=}\Conid{Kleisli}\;(\lambda \Varid{a}\to \Varid{f}\;\Varid{a}\bind \Varid{g}){}\<[E]%

@@

		\subcaptionbox{first\label{t1}}{\includegraphics[width = 1.5in]{images/first}} &
		\subcaptionbox{second\label{fig:secondImg}}{\includegraphics[width = 1.5in]{images/second}} \\
		\subcaptionbox{(***)\label{fig:***Img}}{\includegraphics[width = 1.5in]{images/starstarstar}} &
		\subcaptionbox{(\&\&\&)\label{fig:&&&Img}}{\includegraphics[width = 1.5in]{images/andandand}}\

@@

\>[B]{}\Varid{second}\;\Varid{f}\mathrel{=}\Varid{arr}\;\Varid{swap}\mathbin{>\!>\!>}\Varid{first}\;\Varid{f}\mathbin{>\!>\!>}\Varid{arr}\;\Varid{swap}{}\<[E]%

@@

the \inlinecode{***} combinator which combines \includegraphics{first} and \inlinecode{second} to handle two inputs in one arrow, (Fig.\ref{fig:***Img},~\ref{fig:***})

@@

\>[B]{}(\mathbin{***})\mathbin{::}\Conid{Arrow}\;\Varid{arr}\Rightarrow \Varid{arr}\;\Varid{a}\;\Varid{b}\to \Varid{arr}\;\Varid{c}\;\Varid{d}\to \Varid{arr}\;(\Varid{a},\Varid{c})\;(\Varid{b},\Varid{d}){}\<[E]%

@@

\>[B]{}\Varid{f}\mathbin{***}\Varid{g}\mathrel{=}\Varid{first}\;\Varid{f}\mathbin{>\!>\!>}\Varid{second}\;\Varid{g}{}\<[E]%

@@

\>[B]{}(\mathbin{\&\&\&})\mathbin{::}\Conid{Arrow}\;\Varid{arr}\Rightarrow \Varid{arr}\;\Varid{a}\;\Varid{b}\to \Varid{arr}\;\Varid{a}\;\Varid{c}\to \Varid{a}\;\Varid{a}\;(\Varid{b},\Varid{c}){}\<[E]%

@@

\>[B]{}\Varid{f}\mathbin{\&\&\&}\Varid{g}\mathrel{=}\Varid{arr}\;(\lambda \Varid{a}\to (\Varid{a},\Varid{a}))\mathbin{>\!>\!>}(\Varid{f}\mathbin{***}\Varid{g}){}\<[E]%

@@

A short example given by Hughes on how to use this is \lstinline{add} over arrows, which can be seen in Fig.~\ref{fig:addArrows}.

@@

\>[B]{}\Varid{add}\;\Varid{f}\;\Varid{g}\mathrel{=}(\Varid{f}\mathbin{\&\&\&}\Varid{g})\mathbin{>\!>\!>}\Varid{arr}\;(\lambda (\Varid{u},\Varid{v})\to \Varid{u}\mathbin{+}\Varid{v}){}\<[E]%

@@

% The benefit of using the \inlinecode{Arrow} typeclass is that any type which is shown to be an arrow can be used in conjunction with this newly created \lstinline{add} combinator. Even though this example is quite simple, the power of the arrow interface immediately is clear: If a type is an arrow, it can automatically used together with every library that works on arrows. Compared to simple monads, this enables us to write code that is more extensible, without touching the internals of the specific arrows.

@@

Translating this into arrow terms gives us a new operator \inlinecode{parEvalN} that lifts a list of arrows \inlinecode{[arr a b]} to a parallel arrow \inlinecode{arr [a] [b]} (Fig.~\ref{fig:parEvalNArrowFn}) (This combinator is similar to our utility function \lstinline{listApp} from Appendix \ref{utilfns}, but does parallel instead of serial evaluation).

@@

\>[17]{}\Varid{listApp}\;\Varid{fs}\mathbin{>\!>\!>}{}\<[E]%

@@

\>[17]{}\Varid{arr}\;(\Varid{withStrategy}\;(\Varid{parList}\;\Varid{rdeepseq}))\mathbin{\&\&\&}\Varid{arr}\;\Varid{id}\mathbin{>\!>\!>}{}\<[E]%

@@

\>[17]{}\Varid{arr}\;(\Varid{withStrategy}\;(\Varid{parList}\;\Varid{strat}))\mathbin{\&\&\&}\Varid{arr}\;\Varid{id}\mathbin{>\!>\!>}{}\<[E]%

@@

We therefore apply each function to its corresponding input value with \inlinecode{listApp} (Fig.~\ref{fig:listApp}) and then fork the computation away with \inlinecode{arr spawnP} inside a \inlinecode{zipWithArr} call. This yields a list \inlinecode{[Par (IVar b)]}, which we then convert into \inlinecode{Par [IVar b]} with \inlinecode{arr sequenceA}. In order to wait for the computation to finish, we map over the \inlinecode{IVar}s inside the ParMonad with \inlinecode{arr (>>= mapM get)}. The result of this operation is a \inlinecode{Par [b]} from which we can finally remove the monad again by running \inlinecode{arr runPar} to get our output of \inlinecode{[b]}.
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{instance}\;(\Conid{NFData}\;\Varid{b},\Conid{ArrowApply}\;\Varid{arr},\Conid{ArrowChoice}\;\Varid{arr})\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Conid{ArrowParallel}\;\Varid{arr}\;\Varid{a}\;\Varid{b}\;\Varid{conf}\;\mathbf{where}{}\<[E]%
\\
\>[9]{}\hsindent{8}{}\<[17]%
\>[17]{}\Varid{parEvalN}\;\anonymous \;\Varid{fs}\mathrel{=}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}(\Varid{arr}\mathbin{\$}\lambda \Varid{as}\to (\Varid{fs},\Varid{as}))\mathbin{>\!>\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{zipWithArr}\;(\Varid{app}\mathbin{>\!>\!>}\Varid{arr}\;\Varid{spawnP})\mathbin{>\!>\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{arr}\;\Varid{sequenceA}\mathbin{>\!>\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{arr}\;(\bind \Varid{mapM}\;\Varid{get})\mathbin{>\!>\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{arr}\;\Varid{runPar}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
 %$ %% formatting
\caption{ArrowParallel instance for the Par Monad backend}
\label{fig:ArrowParallelParMonad}
\end{figure}

\subsubsection{Eden}
For the Multicore and ParMonad implementation we could use general instances of \inlinecode{ArrowParallel} that just require the \inlinecode{ArrowApply} and \inlinecode{ArrowChoice} typeclasses. With Eden this is not the case as we can only spawn a list of functions and we cannot extract simple functions out of arrows. While we could still manage to have only one class in the module by introducing a typeclass \inlinecode{ArrowUnwrap} (Fig.~\ref{fig:ArrowUnwrap}).
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{class}\;(\Conid{Arrow}\;\Varid{arr})\Rightarrow \Conid{ArrowUnwrap}\;\Varid{arr}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;\Varid{a}\;\Varid{b}\to (\Varid{a}\to \Varid{b}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{possible ArrowUnwrap typeclass}
\label{fig:ArrowUnwrap}
\end{figure}
We don't do it here for aesthetic resons, though. For now, we just implement \inlinecode{ArrowParallel} for normal functions (Fig.~\ref{fig:ArrowParallelEdenFns})
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{instance}\;(\Conid{Trans}\;\Varid{a},\Conid{Trans}\;\Varid{b})\Rightarrow \Conid{ArrowParallel}\;(\to )\;\Varid{a}\;\Varid{b}\;\Varid{conf}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\Varid{parEvalN}\;\anonymous \;\Varid{fs}\;\Varid{as}\mathrel{=}\Varid{spawnF}\;\Varid{fs}\;\Varid{as}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{ArrowParallel instance for functions in the Eden backend}
\label{fig:ArrowParallelEdenFns}
\end{figure}
and the Kleisli type (Fig.~\ref{fig:ArrowParallelEdenKleisli}).
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{instance}\;(\Conid{Monad}\;\Varid{m},\Conid{Trans}\;\Varid{a},\Conid{Trans}\;\Varid{b},\Conid{Trans}\;(\Varid{m}\;\Varid{b}))\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Conid{ArrowParallel}\;(\Conid{Kleisli}\;\Varid{m})\;\Varid{a}\;\Varid{b}\;\Varid{conf}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\Varid{parEvalN}\;\Varid{conf}\;\Varid{fs}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}(\Varid{arr}\mathbin{\$}\Varid{parEvalN}\;\Varid{conf}\;(\Varid{map}\;(\lambda (\Conid{Kleisli}\;\Varid{f})\to \Varid{f})\;\Varid{fs}))\mathbin{>\!>\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}(\Conid{Kleisli}\mathbin{\$}\Varid{sequence}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{ArrowParallel instance for the Kleisli type in the Eden backend}
\label{fig:ArrowParallelEdenKleisli}
\end{figure}

@@

\subsubsection{Lazy \inlinecode{parEvalN}}
\begin{figure}[h]
	\includegraphics[scale=0.7]{images/parEvalNLazy}
	\caption{Schematic depiction of parEvalNLazy}
	\label{fig:parEvalNLazyImg}
\end{figure}
The function \inlinecode{parEvalN} is 100\% strict, which means that it fully evaluates all passed arrows. Sometimes this might not be feasible, as it will not work on infinite lists of functions like e.g. \inlinecode{map (arr . (+)) [1..]} or just because we need the arrows evaluated in chunks. \inlinecode{parEvalNLazy} (Fig.~\ref{fig:parEvalNLazyImg},~\ref{fig:parEvalNLazy}) fixes this. It works by first chunking the input from \inlinecode{[a]} to \inlinecode{[[a]]} with the given \inlinecode{ChunkSize} in \inlinecode{arr (chunksOf chunkSize)}. These chunks are then fed into a list \inlinecode{[arr [a] [b]]} of parallel arrows created by feeding chunks of the passed \inlinecode{ChunkSize} into the regular parEvalN by using \inlinecode{listApp} (Fig.~\ref{fig:listApp}). The resulting \inlinecode{[[b]]} is lastly converted into \inlinecode{[b]} with \inlinecode{arr concat}.
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{parEvalNLazy}\mathbin{::}(\Conid{ArrowParallel}\;\Varid{arr}\;\Varid{a}\;\Varid{b}\;\Varid{conf},\Conid{ArrowChoice}\;\Varid{arr},\Conid{ArrowApply}\;\Varid{arr})\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{conf}\to \Conid{ChunkSize}\to [\mskip1.5mu \Varid{arr}\;\Varid{a}\;\Varid{b}\mskip1.5mu]\to (\Varid{arr}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{b}\mskip1.5mu]){}\<[E]%
\\
\>[B]{}\Varid{parEvalNLazy}\;\Varid{conf}\;\Varid{chunkSize}\;\Varid{fs}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;(\Varid{chunksOf}\;\Varid{chunkSize})\mathbin{>\!>\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{listApp}\;\Varid{fchunks}\mathbin{>\!>\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;\Varid{concat}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\mathbf{where}\;\Varid{fchunks}\mathrel{=}\Varid{map}\;(\Varid{parEvalN}\;\Varid{conf})\mathbin{\$}\Varid{chunksOf}\;\Varid{chunkSize}\;\Varid{fs}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
 %$ %% formatting
\caption{Definition of parEvalNLazy}
\label{fig:parEvalNLazy}
\end{figure}

>>>>

%% add a tweak to polycode, similar to literate in lstlistings, but uglier
%% ^^^^ you want this. and similar for ***, +++, etc.
%% a hack for list comprehensions' <-, typeset it as <--
%%%%%format <--        = "\in "
%%%% %format ` = "\`"
%%%% %% works fine without

@@

%% fight for pipepipepipe
\newcommand{\pipepipepipe}{\ensuremath{\mathbin{\mid\!\mid\!\mid}}\xspace}
%%% ^^^^ keep consistent with definition at the top of main.lhs
%%%\newcommand{\pipepipepipe}{\inlinecode{|||}\xspace}

@@

%\subfloat[Arrow class definition][]{%

@@

\parbox{0.49\linewidth}{%

@@

\>[3]{}(\mathbin{>\!\!>\!\!>})\mathbin{::}\Varid{arr}\;\Varid{a}\;\Varid{b}\to \Varid{arr}\;\Varid{b}\;\Varid{c}\to \Varid{arr}\;\Varid{a}\;\Varid{c}{}\<[E]%

@@

\caption{Arrow class definition}
}\qquad
\begin{minipage}{0.49\linewidth}%
%\subfloat[The arrow type class definition on the left with schematic depiction of its combinators on the right][]{%
%	\begin{center}

@@

%	\end{center}
%}%\hfill
\caption{The schematic depiction of \inlinecode{Arrow} combinators}
\end{minipage}

@@

\inlinecode{arr} is used to lift an ordinary function to an arrow type, similarly to the monadic \inlinecode{return}. The \inlinecode{>>>} operator is analogous to the monadic composition  \inlinecode{>>=} and combines two arrows \inlinecode{arr a b} and \inlinecode{arr b c} by "wiring" the outputs of the first to the inputs to the second to get a new arrow \inlinecode{arr a c}. Lastly, the \inlinecode{first} operator  takes the input arrow from \inlinecode{b} to \inlinecode{c} and converts it into an arrow on pairs with the second argument untouched. It allows us to to save input across arrows.

@@

The most prominent instances of this interface are regular functions \inlinecode{(->)} (Fig.~\ref{fig:arrowfn}),

@@

\>[9]{}\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{g}\mathrel{=}\Varid{g}\mathbin{\circ}\Varid{f}{}\<[E]%

@@

\>[9]{}\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{g}\mathrel{=}\Conid{Kleisli}\;(\lambda \Varid{a}\to \Varid{f}\;\Varid{a}\bind \Varid{g}){}\<[E]%

@@

		% \subcaptionbox
{first\label{t1}}{\includegraphics[width = 1.5in]{images/first}} &
		% \subcaptionbox
{second\label{fig:secondImg}}{\includegraphics[width = 1.5in]{images/second}} \\
		% \subcaptionbox
{(***)\label{fig:***Img}}{\includegraphics[width = 1.5in]{images/starstarstar}} &
		% \subcaptionbox
{(\&\&\&)\label{fig:&&&Img}}{\includegraphics[width = 1.5in]{images/andandand}}\

@@

\>[B]{}\Varid{second}\;\Varid{f}\mathrel{=}\Varid{arr}\;\Varid{swap}\mathbin{>\!\!>\!\!>}\Varid{first}\;\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{arr}\;\Varid{swap}{}\<[E]%

@@

the \inlinecode{***} combinator which combines \inlinecode{first} and \inlinecode{second} to handle two inputs in one arrow, (Fig.\ref{fig:***Img},~\ref{fig:***})

@@

\>[B]{}(\mathbin{\*\!\*\!\*})\mathbin{::}\Conid{Arrow}\;\Varid{arr}\Rightarrow \Varid{arr}\;\Varid{a}\;\Varid{b}\to \Varid{arr}\;\Varid{c}\;\Varid{d}\to \Varid{arr}\;(\Varid{a},\Varid{c})\;(\Varid{b},\Varid{d}){}\<[E]%

@@

\>[B]{}\Varid{f}\mathbin{\*\!\*\!\*}\Varid{g}\mathrel{=}\Varid{first}\;\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{second}\;\Varid{g}{}\<[E]%

@@

\>[B]{}(\mathbin{\&\!\&\!\&})\mathbin{::}\Conid{Arrow}\;\Varid{arr}\Rightarrow \Varid{arr}\;\Varid{a}\;\Varid{b}\to \Varid{arr}\;\Varid{a}\;\Varid{c}\to \Varid{a}\;\Varid{a}\;(\Varid{b},\Varid{c}){}\<[E]%

@@

\>[B]{}\Varid{f}\mathbin{\&\!\&\!\&}\Varid{g}\mathrel{=}\Varid{arr}\;(\lambda \Varid{a}\to (\Varid{a},\Varid{a}))\mathbin{>\!\!>\!\!>}(\Varid{f}\mathbin{\*\!\*\!\*}\Varid{g}){}\<[E]%

@@

A short example given by Hughes on how to use this is \inlinecode{add} over arrows, which can be seen in Fig.~\ref{fig:addArrows}.

@@

\>[B]{}\Varid{add}\;\Varid{f}\;\Varid{g}\mathrel{=}(\Varid{f}\mathbin{\&\!\&\!\&}\Varid{g})\mathbin{>\!\!>\!\!>}\Varid{arr}\;(\lambda (\Varid{u},\Varid{v})\to \Varid{u}\mathbin{+}\Varid{v}){}\<[E]%

@@

% The benefit of using the \inlinecode{Arrow} typeclass is that any type which is shown to be an arrow can be used in conjunction with this newly created \inlinecode{add} combinator. Even though this example is quite simple, the power of the arrow interface immediately is clear: If a type is an arrow, it can automatically used together with every library that works on arrows. Compared to simple monads, this enables us to write code that is more extensible, without touching the internals of the specific arrows.

@@

Translating this into arrow terms gives us a new operator \inlinecode{parEvalN} that lifts a list of arrows \inlinecode{[arr a b]} to a parallel arrow \inlinecode{arr [a] [b]} (Fig.~\ref{fig:parEvalNArrowFn}) (This combinator is similar to our utility function \inlinecode{listApp} from Appendix~\ref{utilfns}, but does parallel instead of serial evaluation).

@@

\>[17]{}\Varid{listApp}\;\Varid{fs}\mathbin{>\!\!>\!\!>}{}\<[E]%

@@

\>[17]{}\Varid{arr}\;(\Varid{withStrategy}\;(\Varid{parList}\;\Varid{rdeepseq}))\mathbin{\&\!\&\!\&}\Varid{arr}\;\Varid{id}\mathbin{>\!\!>\!\!>}{}\<[E]%

@@

\>[17]{}\Varid{arr}\;(\Varid{withStrategy}\;(\Varid{parList}\;\Varid{strat}))\mathbin{\&\!\&\!\&}\Varid{arr}\;\Varid{id}\mathbin{>\!\!>\!\!>}{}\<[E]%

@@

We therefore apply each function to its corresponding input value with \inlinecode{listApp} (Fig.~\ref{fig:listApp}) and then fork the computation away with \inlinecode{arr spawnP} inside a \inlinecode{zipWithArr} call. This yields a list \inlinecode{[Par (IVar b)]}, which we then convert into \inlinecode{Par [IVar b]} with \inlinecode{arr sequenceA}. In order to wait for the computation to finish, we map over the \inlinecode{IVar}s inside the ParMonad with \inlinecode{arr (>>= mapM get)}. The result of this operation is a \inlinecode{Par [b]} from which we can finally remove the monad again by running \inlinecode{arr runPar} to get our output of \inlinecode{[b]}.
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{instance}\;(\Conid{NFData}\;\Varid{b},\Conid{ArrowApply}\;\Varid{arr},\Conid{ArrowChoice}\;\Varid{arr})\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Conid{ArrowParallel}\;\Varid{arr}\;\Varid{a}\;\Varid{b}\;\Varid{conf}\;\mathbf{where}{}\<[E]%
\\
\>[9]{}\hsindent{8}{}\<[17]%
\>[17]{}\Varid{parEvalN}\;\anonymous \;\Varid{fs}\mathrel{=}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}(\Varid{arr}\mathbin{\$}\lambda \Varid{as}\to (\Varid{fs},\Varid{as}))\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{zipWithArr}\;(\Varid{app}\mathbin{>\!\!>\!\!>}\Varid{arr}\;\Varid{spawnP})\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{arr}\;\Varid{sequenceA}\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{arr}\;(\bind \Varid{mapM}\;\Varid{get})\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{arr}\;\Varid{runPar}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
 %$ %% formatting
\caption{ArrowParallel instance for the Par Monad backend}
\label{fig:ArrowParallelParMonad}
\end{figure}

\subsubsection{Eden}
For the Multicore and ParMonad implementation we could use general instances of \inlinecode{ArrowParallel} that just require the \inlinecode{ArrowApply} and \inlinecode{ArrowChoice} typeclasses. With Eden this is not the case as we can only spawn a list of functions and we cannot extract simple functions out of arrows. While we could still manage to have only one class in the module by introducing a typeclass \inlinecode{ArrowUnwrap} (Fig.~\ref{fig:ArrowUnwrap}).
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{class}\;(\Conid{Arrow}\;\Varid{arr})\Rightarrow \Conid{ArrowUnwrap}\;\Varid{arr}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;\Varid{a}\;\Varid{b}\to (\Varid{a}\to \Varid{b}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{possible ArrowUnwrap typeclass}
\label{fig:ArrowUnwrap}
\end{figure}
We don't do it here for aesthetic resons, though. For now, we just implement \inlinecode{ArrowParallel} for normal functions (Fig.~\ref{fig:ArrowParallelEdenFns})
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{instance}\;(\Conid{Trans}\;\Varid{a},\Conid{Trans}\;\Varid{b})\Rightarrow \Conid{ArrowParallel}\;(\to )\;\Varid{a}\;\Varid{b}\;\Varid{conf}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\Varid{parEvalN}\;\anonymous \;\Varid{fs}\;\Varid{as}\mathrel{=}\Varid{spawnF}\;\Varid{fs}\;\Varid{as}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{ArrowParallel instance for functions in the Eden backend}
\label{fig:ArrowParallelEdenFns}
\end{figure}
and the Kleisli type (Fig.~\ref{fig:ArrowParallelEdenKleisli}).
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{instance}\;(\Conid{Monad}\;\Varid{m},\Conid{Trans}\;\Varid{a},\Conid{Trans}\;\Varid{b},\Conid{Trans}\;(\Varid{m}\;\Varid{b}))\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Conid{ArrowParallel}\;(\Conid{Kleisli}\;\Varid{m})\;\Varid{a}\;\Varid{b}\;\Varid{conf}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\Varid{parEvalN}\;\Varid{conf}\;\Varid{fs}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}(\Varid{arr}\mathbin{\$}\Varid{parEvalN}\;\Varid{conf}\;(\Varid{map}\;(\lambda (\Conid{Kleisli}\;\Varid{f})\to \Varid{f})\;\Varid{fs}))\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}(\Conid{Kleisli}\mathbin{\$}\Varid{sequence}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{ArrowParallel instance for the Kleisli type in the Eden backend}
\label{fig:ArrowParallelEdenKleisli}
\end{figure}

@@

\subsubsection{Lazy \inlinecode{parEvalN}}
\begin{figure}[h]
	\includegraphics[scale=0.7]{images/parEvalNLazy}
	\caption{Schematic depiction of parEvalNLazy}
	\label{fig:parEvalNLazyImg}
\end{figure}
The function \inlinecode{parEvalN} is 100\% strict, which means that it fully evaluates all passed arrows. Sometimes this might not be feasible, as it will not work on infinite lists of functions like e.g. \inlinecode{map (arr . (+)) [1..]} or just because we need the arrows evaluated in chunks. \inlinecode{parEvalNLazy} (Fig.~\ref{fig:parEvalNLazyImg},~\ref{fig:parEvalNLazy}) fixes this. It works by first chunking the input from \inlinecode{[a]} to \inlinecode{[[a]]} with the given \inlinecode{ChunkSize} in \inlinecode{arr (chunksOf chunkSize)}. These chunks are then fed into a list \inlinecode{[arr [a] [b]]} of parallel arrows created by feeding chunks of the passed \inlinecode{ChunkSize} into the regular parEvalN by using \inlinecode{listApp} (Fig.~\ref{fig:listApp}). The resulting \inlinecode{[[b]]} is lastly converted into \inlinecode{[b]} with \inlinecode{arr concat}.
\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{parEvalNLazy}\mathbin{::}(\Conid{ArrowParallel}\;\Varid{arr}\;\Varid{a}\;\Varid{b}\;\Varid{conf},\Conid{ArrowChoice}\;\Varid{arr},\Conid{ArrowApply}\;\Varid{arr})\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{conf}\to \Conid{ChunkSize}\to [\mskip1.5mu \Varid{arr}\;\Varid{a}\;\Varid{b}\mskip1.5mu]\to (\Varid{arr}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{b}\mskip1.5mu]){}\<[E]%
\\
\>[B]{}\Varid{parEvalNLazy}\;\Varid{conf}\;\Varid{chunkSize}\;\Varid{fs}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;(\Varid{chunksOf}\;\Varid{chunkSize})\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{listApp}\;\Varid{fchunks}\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;\Varid{concat}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\mathbf{where}\;\Varid{fchunks}\mathrel{=}\Varid{map}\;(\Varid{parEvalN}\;\Varid{conf})\mathbin{\$}\Varid{chunksOf}\;\Varid{chunkSize}\;\Varid{fs}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
 %$ %% formatting
\caption{Definition of parEvalNLazy}
\label{fig:parEvalNLazy}
\end{figure}
